---
title: "Critical Result Callback Monitor"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill

---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(xts)
library(dygraphs)
library(plotly)
library(highcharter)
library(kableExtra)
library(DT)
```

```{r data, include = FALSE}
cb_data <- read_csv("data/july_critical_callback_data.csv") %>%
            mutate(call_hour = hour(last_phone_datetime),
                   call_wday = wday(last_phone_datetime, 
                                    label = TRUE), 
                   call_date = day(last_phone_datetime),
                   call_week = isoweek(last_phone_datetime),
                   call_month = month(last_phone_datetime),
                   call_year = year(last_phone_datetime),
                   test_code = replace_na(test_code, "NA"))
```

Overall {data-icon="fa-phone-volume"}
=============================

Row
--------------------------------------
### Average TAT
```{r avg-tat}

mean_tat <- round(mean(cb_data$call_tat), 0)
valueBox(value = mean_tat, icon = "fa-stopwatch", 
         caption = "Mean Callback Time", color = "#708090")
```

### Total calls
```{r total-calls}
calls <- cb_data %>% nrow()
valueBox(value = calls, icon = "fa-hashtag", 
         caption = "Total calls", color = "orange")

```

### Percent on Time
```{r pct-on-time}
ontime_n <- cb_data %>% filter(call_tat <= 30) %>% nrow()
late_n <- cb_data %>% filter(call_tat >30) %>% nrow()
pct_ontime <- round((ontime_n/calls)*100,0)
#pct_ontime <- scales::label_percent()(ontime_n/calls)

if(pct_ontime >= 95){
  valueBox(value = pct_ontime, icon = "fa-percent", 
           caption = "Percent on time", color = "mediumseagreen") #3CB371
} else{
valueBox(value = pct_ontime, icon = "fa-percent", 
         caption = "Percent on time", color = "indianred")} #CD5C5C
```


Row
----------------------------------

### Weekly Overview: Hourly Call Breakdown for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r plot-hourly-by-day-no-weekend}

p1 <- ggplot(filter(cb_data, 
              call_wday %in% c("Mon", "Tue", "Wed", 
                               "Thu", "Fri"))) +
  geom_bar(aes(x = call_hour, fill = tech_location)) + 
  facet_grid(call_wday~call_week) +
  labs(x = "Hour", y = "Count", fill = "Type") +
  scale_fill_manual(values=c("steelblue3", "gray60")) +
  theme_bw() +
  theme(legend.position = "top")

ggplotly(p1) %>% layout(legend = list(orientation = "h", 
                                      x = 0.35, y = 1.2)) 
# have to manually position the legend

```

### Daily Call Volumes for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r plot-daily-volumes}

daily_vol <- cb_data %>% 
              group_by(call_year, call_month, call_date) %>%
              summarize(n = n()) %>%
              ungroup() %>%
              unite(dttm, call_year, call_month, call_date, 
                    sep = "-", remove = TRUE) %>%
              mutate(dttm = as.Date(dttm))

daily_vol_ts <- xts(daily_vol$n, order.by = daily_vol$dttm)

p2 <- dygraph(daily_vol_ts) %>% 
  dyOptions(colors = "#4F94CD", strokeWidth = 2) 
p2
```


Call Center {data-icon="fa-users"}
=============================

Row
--------------------------------------
### Average TAT
```{r avg-tat-cc}
cc_calls <- cb_data %>% filter(tech_location == "CallCenter")
mean_tat_cc <- round(mean(cc_calls$call_tat),0)
 
valueBox(value = mean_tat_cc, icon = "fa-stopwatch", 
         caption = "Mean Callback Time", color = "#708090")
```

### Total calls
```{r total-calls-cc}
cc_calls_n <- cb_data %>% filter(tech_location == "CallCenter") %>% nrow()
valueBox(value = cc_calls_n, icon = "fa-hashtag", 
         caption = "Total calls", color = "orange")

```

### Percent on Time
```{r pct-on-time-cc}
cc_ontime_n <- cb_data %>% 
  filter(tech_location == "CallCenter", call_tat <= 30) %>% nrow()
cc_late_n <- cb_data %>% 
  filter(tech_location == "CallCenter", call_tat >30) %>% nrow()
cc_pct_ontime <- round((cc_ontime_n/cc_calls_n)*100,0)
#pct_ontime <- scales::label_percent()(ontime_n/calls)

if(cc_pct_ontime >= 95){
  valueBox(value = cc_pct_ontime, icon = "fa-percent", 
           caption = "Percent on time", color = "mediumseagreen") #3CB371
} else{
valueBox(value = cc_pct_ontime, icon = "fa-percent", 
         caption = "Percent on time", color = "indianred")} #CD5C5C
```

Row 
--------------------------------------

### Overdue Call Details for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r cc-overdue}
cc_late <- cb_data %>% 
  filter(tech_location == "CallCenter", call_tat >30) %>%
  select(call_tat, accession, pt_type, pt_loc_code, test_code,
         result_datetime, phoned_title, tech)

datatable(cc_late, options = list(pageLength = 20, autoWidth = TRUE))

```


### Tech Summary for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r cc-tech}
p3 <- hcboxplot(
  outliers = FALSE,
  x = cc_calls$call_tat,
  var = cc_calls$tech,
  name = "TAT", 
  color = "#4F94CD",
  lineWidth = 2) %>%
  hc_title(text = "") %>%
  hc_xAxis(title = list(text = "Tech Code")) %>%
  hc_yAxis(title = list(text = "Call Time in Min")) %>%
  hc_chart(type = "column")
p3
```

Non Call Center {data-icon="fa-users"}
=============================

Row 
--------------------------------------
### Average TAT
```{r avg-tat-non}
non_calls <- cb_data %>% filter(tech_location == "Non-CC")
mean_tat_non <- round(mean(non_calls$call_tat),0)
 
valueBox(value = mean_tat_non, icon = "fa-stopwatch",
         caption = "Mean Callback Time", color = "#708090")
```

### Total calls
```{r total-calls-non}
non_calls_n <- cb_data %>% filter(tech_location == "Non-CC") %>% nrow()
valueBox(value = non_calls_n, icon = "fa-hashtag",
         caption = "Total calls", color = "orange")

```

### Percent on Time
```{r pct-on-time-non}
non_ontime_n <- cb_data %>% 
  filter(tech_location == "Non-CC", call_tat <= 30) %>% nrow()
non_late_n <- cb_data %>% 
  filter(tech_location == "Non-CC", call_tat >30) %>% nrow()
non_pct_ontime <- round((non_ontime_n/non_calls_n)*100,0)
#pct_ontime <- scales::label_percent()(ontime_n/calls)

if(non_pct_ontime >= 95){
  valueBox(value = non_pct_ontime, icon = "fa-percent", 
           caption = "Percent on time", color = "mediumseagreen") #3CB371
} else{
valueBox(value = non_pct_ontime, icon = "fa-percent", 
         caption = "Percent on time", color = "indianred")} #CD5C5C
```

Row 
--------------------------------------

### Overdue Call Details for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r non-overdue}
non_late <- cb_data %>% 
  filter(tech_location == "Non-CC", call_tat >30) %>%
  select(call_tat, accession, pt_type, pt_loc_code, test_code,
         result_datetime, phoned_title, tech)

datatable(non_late, options = list(pageLength = 20, autoWidth = TRUE))

```


### Tech Summary for `r date(min(cb_data$last_phone_datetime))` through `r date(max(cb_data$last_phone_datetime))`
```{r non-tech}
p4 <- hcboxplot(
  outliers = FALSE,
  x = non_calls$call_tat,
  var = non_calls$tech,
  name = "TAT", 
  color = "#4F94CD",
  lineWidth = 2) %>%
  hc_title(text = "") %>%
  hc_xAxis(title = list(text = "Tech Code")) %>%
  hc_yAxis(title = list(text = "Call Time in Min")) %>%
  hc_chart(type = "column")
p4
```

